Should I Airbnb My Property?

### 2017-03-21 abb_Writeup.Rmd
Final edits

@@

As our scenario compares a hypothetical and particular owner over the course of a single  year the tax $T$ and $F$ costs are identical between the two leasing strategy and therefore cancel out. General maintenance $M$ expenses may differ between the two strategies, and an argument could be made for either to be higher. Long-term renters spend more time in the property over the course of the year and bring their own appliances^[White goods are usually not supplied in the Australian rental market.] and furniture which may damage the home. They are also more likely to affix items to the structure.  Conversely, short-term rentals are usually furnished and the occupants have less incentive not to be evicted as their damage is not often uncovered until after they have left^[Note that Airbnb's rating policy may help protect hosts as bad guests can be flagged on the platform.].  Both arguments are meritous and we were unable to find any evidence to support maintenance costs being higher in short- or long-term rentals. As a result we have made the assumption that the maintenance costs will be identical, and they too cancel out.

@@

Janitorial costs, $J$, cover the costs to 'turn-over' the property between occupants.  For a short-term leasing strategy, these costs can be experienced every day or two, depending on the minimum length of stay and the total number of occupancies over the course of the year. Our data for short-term rentals comes from the Airbnb platform.  In addition to the nightly rate that an occupant pays to the owner there is also a separate 'cleaning fee' that is attached, as is a booking fee to Airbnb^[This fee could be considered analogous to the leasing fees $L$ that long-term rentals pay]. Some hosts add additional 'linen' fees to cover costs of laundering.  While it may be advantageous for hosts in competitive markets to underquote their cleaning fees to keep occupancies up, we have no evidence to suggest that owners are exhibiting such behavior. As a result of janitorial fees being a separate line item and directly paid by the occupants they can also be ignored in the comparative analysis below as they are direct.

@@

While this simple analysis does not control for location and product differences, it does highlight the fact that some Airbnb owners are likely making greater revenues than in long-term rentals, but that most are likely not.  When looking at potential revenues, however, a great deal more may have the potential to produce additional revenue.  The high number of very low yield Airbnb properties in the Actual condition suggests that some owners may not be focused on maximizing revenue, but instead prefer the flexibility of short-term leasing^[Do note that we have not considered any overall market demand issues when analyzing potential revenue.].

@@

Along with occupancy rates, the other important consideration is the nightly rate.  High occupancy rates with low nightly rates may not be any better than lower occupancy rate with a higher nightly rate.  In Figure 10 (left panel) we plot nightly rates versus occupancy rates.  We then divide the area of the scatter plot into 400 equal areas (20 x 20) and color each based on which is more common in the area, short-term or long-term revenue.  The transparency is related to the total count of actual market properties falling into that area.

@@

Comparing annual revenues from the two leasing strategies shows that revenues from more traditional long-term rentals are, on average, higher and much less variable than those from the Airbnb market (\autoref{revdens}).  However, there is a long right tail to the distribution of short-term revenues suggesting that some owners are doing better, and occasionally much better, in the short-term market than they would on the long-term market. This univariate comparison of revenues does not tell the entire story, as the Airbnb market does not have the same spatial distribution as the long-term market (Figure 8).  Short-term leasing is much more common in the central areas of the city, with small clusters showing up in suburban beach locations and rural tourist destinations like the Yarra Valley wine district.

>>>>

  abb.col[1] <- abb.col[3]
  abb.col[5] <- abb.col[6]

@@

As our scenario compares a hypothetical and particular owner over the course of a single  year the tax $T$ and $F$ costs are identical between the two leasing strategy and therefore cancel out. General maintenance $M$ expenses may differ between the two strategies, and an argument could be made for either to be higher. Long-term renters spend more time in the property over the course of the year and bring their own appliances^[White goods are usually not supplied in the Australian rental market.] and furniture which may damage the home. They are also more likely to affix items to the structure.  Conversely, short-term rentals are usually furnished and the occupants have less incentive not to be evicted as their damage is not often uncovered until after they have left.^[Note that Airbnb's rating policy may help protect hosts as bad guests can be flagged on the platform.]  Both arguments are meritous and we were unable to find any evidence to support maintenance costs being higher in short- or long-term rentals. As a result we have made the assumption that the maintenance costs will be identical, and they too cancel out.

@@

Janitorial costs, $J$, cover the costs to 'turn-over' the property between occupants.  For a short-term leasing strategy, these costs can be experienced every day or two, depending on the minimum length of stay and the total number of occupancies over the course of the year. Our data for short-term rentals comes from the Airbnb platform.  In addition to the nightly rate that an occupant pays to the owner there is also a separate 'cleaning fee' that is attached, as is a booking fee to Airbnb.^[This fee could be considered analogous to the leasing fees $L$ that long-term rentals pay.] Some hosts add additional 'linen' fees to cover costs of laundering.  While it may be advantageous for hosts in competitive markets to underquote their cleaning fees to keep occupancies up, we have no evidence to suggest that owners are exhibiting such behavior. As a result of janitorial fees being a separate line item and directly paid by the occupants they can also be ignored in the comparative analysis below as they are direct.

@@

While this simple analysis does not control for location and product differences, it does highlight the fact that some Airbnb owners are likely making greater revenues than in long-term rentals, but that most are likely not.  When looking at potential revenues, however, a great deal more may have the potential to produce additional revenue.  The high number of very low yield Airbnb properties in the Actual condition suggests that some owners may not be focused on maximizing revenue, but instead prefer the flexibility of short-term leasing.^[Do note that we have not considered any overall market demand issues when analyzing potential revenue.]

@@

Along with occupancy rates, the other important consideration is the nightly rate.  High occupancy rates with low nightly rates may not be any better than lower occupancy rate with a higher nightly rate.  In \autoref{metrohm} (left panel) we plot nightly rates versus occupancy rates.  We then divide the area of the scatter plot into 400 equal areas (20 x 20) and color each based on which is more common in the area, short-term or long-term revenue.  The transparency is related to the total count of actual market properties falling into that area.

@@

Comparing annual revenues from the two leasing strategies shows that revenues from more traditional long-term rentals are, on average, higher and much less variable than those from the Airbnb market (\autoref{revdens}).  However, there is a long right tail to the distribution of short-term revenues suggesting that some owners are doing better, and occasionally much better, in the short-term market than they would on the long-term market. This univariate comparison of revenues does not tell the entire story, as the Airbnb market does not have the same spatial distribution as the long-term market (\autoref{pointmaps}).  Short-term leasing is much more common in the central areas of the city, with small clusters showing up in suburban beach locations and rural tourist destinations like the Yarra Valley wine district.
